\sffamily An EEG-based visual Brain-Computer Interface for gaze-free communication

### 2024-09-22 chapters/bci.tex
Decoders section

Usually, a decoder makes a prediction for a given \emph{trial} while operating
the \ac{bci}.
A trial is the smallest unit on which a selection decision can be made, for
instance be one imagined movement, or one repetition of flashing all targets
in a visual \ac{erp} \ac{bci}.
Some valuable metrics to asses decoder performance in the classification case
are accuracy,
concept of trial
accuracy

single erp responses: roc auc, higher roc auc usually results in higher
accuracy

Finall, an important concept in the evaluation of a \ac{bci} decoder, and of \acp{bci}
in general, is \ac{itr}.
\Ac{itr} reflects how fluently a \ac{bci} can be used for communication and can
be calculated as
\begin{equation}
	\text{ITR} = Q\left(\log_2N+P\log_2P+(1-P)\log_2\frac{1-P}{N-1}\right)
\end{equation}
\Ac{itr} is expressed in bits/s and is dependent on $N$, the number of
different options that can be selected per trial, $P$, the selection accuracy
of the decoder, and $Q$, the number of trials per second.
The parameters $N$, $P$ and $Q$ of this formula give us some insight in the
building blocks a successful, high-\ac{itr} \ac{bci}.
To improve performance, we can aim to
\todo{enumerate}
\begin{enumerate*}[label={\arabic*})]
\item increase $N$ by selecting a paradigm and interface that offer a
broad range of informative selections per trial,
\item increase $P$ by engineering more performant machine learning methods for
decoding, and
\item increase $Q$ by selecting a paradigm and allows fast stimulation or
responses.
\end{enumerate*}

>>>>

Usually, a decoder makes a prediction for a given \emph{trial} while operating
the \ac{bci}.
A trial is the smallest unit on which a selection decision can be made, for
instance be one imagined movement, or one repetition of flashing all targets
in a visual \ac{erp} \ac{bci}.
Some valuable metrics to asses decoder performance in the classification case
are accuracy and \ac{rocauc}.
Accuracy can be obtained from the correctness of predicted trial selection as
\begin{equation}
  \text{accuracy} = \frac{\text{TP}+\text{TN}}{\text{TP}+\text{TN}+\text{FP}+\text{FN}}
\end{equation}
with TP and TN the numbers of true positives and negatives and FP and FP the
numbers of false positives and negatives.
\Ac{rocauc} is also a measure of classifier predictive power, but is more
suited for evaluation of classification of single epochs of data and in the
presence of imbalanced data, e.g. when
comparing single \ac{erps}.
Higer \ac{rocauc} usually translates to a higher target selection accuracy.

Finally, an important concept in the evaluation of a \ac{bci} decoder, and of \acp{bci}
in general, is \ac{itr}.
\Ac{itr} reflects how fluently a \ac{bci} can be used for communication and can
be calculated as
\begin{equation}
	\text{ITR} = Q\left(\log_2N+P\log_2P+(1-P)\log_2\frac{1-P}{N-1}\right)
\end{equation}
\Ac{itr} is expressed in bits/s and is dependent on $N$, the number of
different options that can be selected per trial, $P$, the selection accuracy
of the decoder, and $Q$, the number of trials per second.
The parameters $N$, $P$ and $Q$ of this formula give us some insight in the
building blocks a successful, high-\ac{itr} \ac{bci}.
To improve performance, we can aim to
\todo{enumerate}
\begin{enumerate*}[label={\arabic*})]
\item increase $N$ by selecting a paradigm and interface that offer a
broad range of informative selections per trial,
\item increase $P$ by engineering more performant machine learning methods for
decoding, and
\item increase $Q$ by selecting a paradigm and allows fast stimulation or
responses.
\end{enumerate*}

### 2024-09-22 chapters/gaze_independence.tex
Decoders section

\emph{Parts oare dissociated in the
Section~\ref{sec:gaze-independence} have been published in
\textcite{VanDenKerchove2024}.}

@@

\todo{Regularized spatiotemporal beamforming}
Due to the lack of the N2 component and decrease of amplitude of the
P3 component in covert attention
settings~\cite{Treder2010}, the signal-to-noise ratio (SNR) of
the ERP is lower than in overt attention settings.
Therefore, a straightforward way to reach satisfactory
gaze-independent decoding performance, might be by increasing overall ERP
decoding performance. A more accurate classifier could yield relative
improvements in those settings where performance is not yet near the achievable
maximum.

@@

\todo{Classifier-based Latency Estimation with Woody iterations}

@@

\todo{Tensor discriminant analysis}

@@

\todo{Healthy participant study}
In a first series of experiments, we recorded data with this interface from healthy participants.
The goal of these experiments is to benchmark
gaze-independent ERP decoding algorithms.
We denote this dataset as the Covert Visuospatial Attention ERP dataset
(CVSA-ERP)
This study was approved by the Ethics Commission of University Hospital Leuven
(S62547).
To simulate the dissociation between the eye gaze (visual attention) and the
intended target (mental attention) which could occur in patients with hampered eye
motor control, healthy participants are cued to operate the BCI in specific
\ac{vsa} conditions.

\todo{mention multi-center patient study case studies in cooperation
After data collection from healthy participants and decoder development, we
will carry out a patient study using a similar interface.
The goal of the patient study is to evaluate the influence of eye motor
deficits in patient populations on existing state-of-the-art decoders and on
our proposed decoding methods.
}
\todo{update with current results}

>>>>

\emph{Parts of Section~\ref{sec:gaze-independence} have been published in \textcite{VanDenKerchove2024}.}

@@

\subsubsection{Regularized spatiotemporal beamforming}
Due to the lack of the N2 component and decrease of amplitude of the
P3 component in covert attention
settings~\cite{Treder2010}, the signal-to-noise ratio (SNR) of
the ERP is lower than in overt attention settings.
Therefore, a straightforward way to reach satisfactory
gaze-independent decoding performance, might be by increasing overall ERP
decoding performance. A more accurate classifier could yield relative
improvements in those settings where performance is not yet near the achievable
maximum.

@@

\subsubsection{Classifier-based Latency Estimation with Woody iterations}

@@

\subsubsection{Tensor discriminant analysis}

@@

\subsubsection{Healthy control study}
In a first series of experiments, we recorded data with this interface from healthy participants.
The goal of these experiments is to benchmark
gaze-independent ERP decoding algorithms.
We denote this dataset as the Covert Visuospatial Attention ERP dataset
(CVSA-ERP)
This study was approved by the Ethics Commission of University Hospital Leuven
(S62547).
To simulate the dissociation between the eye gaze (visual attention) and the
intended target (mental attention) which could occur in patients with hampered eye
motor control, healthy participants are cued to operate the BCI in specific
\ac{vsa} conditions.

\subsubsection{Patient case study}
\todo{mention multi-center patient study case studies in cooperation
After data collection from healthy participants and decoder development, we
will carry out a patient study using a similar interface.
The goal of the patient study is to evaluate the influence of eye motor
deficits in patient populations on existing state-of-the-art decoders and on
our proposed decoding methods.
}
\todo{update with current results}
